---
title: 'Bootstrap Analysis'
author: "Vanessa Rathbone"
date: "3/6/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(boot)
library(gt)
library(patchwork)
library(broom)

```


**Study information:** In the study, 22 out of 36 surveyed UCSB community members (61%) identifying as nonbinary/genderqueer responded that they had personally experienced “exclusionary, offensive, hostile or intimidating conduct” (compared to 26% and 19% for those identifying as women and men, respectively). 

**Data source:** The 2014 UCSB Campus Climate Project Final Report (prepared by Rankin & Associates Consulting, available at http://campusclimate.ucop.edu/_common/files/pdf-climate/ucsb-full-report.pdf).

## Create vector of the data
```{r}

# Make a vector with 36 sample size
ucsb_sample <- data.frame("gender_identity" = "nonbinary",
                        "exclusion" = c(1, 1, 1, 1, 1,
                                        1, 1, 1, 1, 1,
                                        1, 1, 1, 1, 1,
                                        1, 1, 1, 1, 1,
                                        1, 1, 0, 0 ,0, 0,
                                        0, 0, 0, 0, 0,
                                        0, 0, 0, 0, 0))

# Vizualize the data
hist(ucsb_sample$exclusion)
mean(ucsb_sample$exclusion)
t.test(ucsb_sample$exclusion) # Get 95% CI for t-distribution

```


## Create a function to find the proportion
```{r}

# Build a function for the proportion of nonbinary/genderqueer individuals experiencing exclusionary conduct based on the 2014 UCSB survey data.
prop_fun <- function(x,i) {sum(x[i])/length(x[i])}

prop_vec <- ucsb_sample$exclusion

ucsbboot_10000 <- boot(prop_vec, 
                    statistic = prop_fun,
                    R = 10000)

```


## Histogram of 10,000 sample bootstrap
```{r}

# These are all the proportions for each of the 10000 bootstrap samples
ucsb_10000_df <- data.frame(bs_proportion = ucsbboot_10000$t) 

ggplot(ucsb_10000_df, aes(bs_proportion)) +
  geom_histogram(bins = 25, color = "brown3", fill = "brown3") +
  theme_minimal() +
  labs(x = "Proportion nonbinary/genderqueer students to experience exclusionary conduct",
       y = "Count of samples from bootstrapping")

```

## Confidence interval
```{r}

boot.ci(ucsbboot_10000, conf = 0.95)

```

Here, the intervals are based on a sampling distribution of means calculated from a large number of bootstrap samples (10,000). It uses the middle 95% of the bootstrap estimates (removing the bottom/top 2.5%) to for the 95% confidence interval. That means we expect that 95 of 100 calculated confidence intervals will contain the actual population parameter. 

The mean proportion of UCSB students that identify as nonbinary/genderqueer who responded that they have personally experienced "exclusionary, offensive, hostile or intimidating conduct is 61% from the original sample size of n = 36. Whereas, with a bootstrapped 95% confidence interval the mean proportion increase to (45.06%, 77.24%) with a bootstrap sample size of n = 10,000.

These results are bootstrap variances needed for studentized intervals, which are used when the statistic is homoscedastic and the standard error of the bootstrap statistic can be estimated by second-stage resampling. Here, the normal intervals are (0.4506,  0.7724) so we can be 95% certain that the actual correlation between nonbinary students experiencing exclusionary conduct in this interval is (45.06%, 77.24%). Because there is no "type" of parameter applied to this function, the interval changes for each type of computation, Normal, Basic, Percentile and BCa. 