---
title: "Parameter Estimation - Wild Fish Catch"
author: "Vanessa Rathbone"
date: "3/5/2021"
output:
    html_document:
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(broom)
library(nlstools)
library(here)
library(janitor)
library(kableExtra)

```

Nonlinear least squares (NLS) finds parameter estimates to minimize the sum of squares of residuals, using an iterative algorithm. We'll use NLS to find the parameters for a logistic growth equation. 

**Data Source:** Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), [Global Capture Production and Global Aquaculture Production, electronic databases](www.fao.org/fishery/topic/16140/en)

## Read in the data and visualize it
```{r, include = FALSE}

fish_df <- read_csv("fish_catch.csv") %>% 
  janitor::row_to_names(row_number = 2) %>% 
  slice(-1) %>% 
  slice(-1) %>% 
  clean_names() %>% 
  filter(year %in% c(1950:2012)) %>% 
  mutate(across(where(is.character), as.numeric)) %>% 
  mutate(year_number = 0:62)

```

```{r, echo = FALSE}
ggplot(fish_df, aes(x = year, y = wild_catch)) + 
  geom_point()
```

There is a linear relationship between wild fish caught over time. Around the 1990s, however, you can see some type of plateau, either a threshold of fish being caught due to abundance (or carrying capacity), or a limit to how many fish can be caught due to permitting. This can be described mathematically by the following logistic growth equation: 

$P(t)=\frac{K}{1+Ae^{-kt}}$, where

- $K$ is the carrying capacity
- $A$ is $\frac{K-P_0}{P_0}$
- $P_0$ is the population at time zero
- $P(t)$ is the population at time $t$
- $k$ is the growth rate constant

```{r, echo = FALSE}
# Find linear coefficients
# Get up to 38 years and log transform the wild catch
# We do this so we can estimate the growth rate constant (k) *during exponential growth phase)
df_lm <- fish_df %>% 
  filter(year_number < 38) %>% 
  mutate(ln_wild = log(wild_catch))

lm_fish <- lm(ln_wild ~ year_number, data = df_lm)

A <- (90 - 15)/15
  
# Model linear to get *k* estimate (the slope of this linear equation is an estimate of the growth rate constant):
lm_k <- lm(ln_wild ~ year_number, data = df_lm)

# Coefficient (k) ~ 0.04

```

Initial estimates for the parameters in the model are:

- $K$ is ~ 90
- $A$ is ~ 5
- $k$ is ~ 0.04

## Non-Linear Least Squares Calculation
```{r, echo=FALSE, results='hide'}

df_nls <- nls(wild_catch ~ K/(1 + A*exp(-r*year_number)),
              data = fish_df,
              start = list(K = 90, A = 15, r = 0.04),
              trace = TRUE
              )

#prep data for table 
fish_table <- broom::tidy(df_nls) %>% 
  mutate_if(is.numeric, funs(as.character(signif(., 3))))

```

**Table 1.** Non-Linear Least Squares for Wild Fish Caught (million tons) 1950-2016.
```{r, echo = FALSE}

fish_table %>% 
  kable(col.names = c("Variable",
                      "Estimated Value", 
                      "Standard Error",
                      "Statistic",
                      "P-value")) %>% 
# Column names - can't change row names in kable.
  kable_styling(bootstrap_options = "striped",
                full_width = F,
                position = "left") %>% 
  add_header_above(bold = TRUE,
                   line = TRUE, 
                   c("Non-Linear Least Squares for Wild Fish Caught (million tons) 1950-2016" = 5))
  
```
## Finalized visualization of observed vs. predicted values
```{r, echo = FALSE}

fish_predict <- predict(df_nls)

fish_complete <- data.frame(fish_df, fish_predict)

ggplot(data = fish_complete, aes(x = year, y = wild_catch)) +
  geom_point(aes(colour = "Observed")) +
  geom_line(aes(x = year, 
                y = fish_predict,
                colour = "Predicted")) +
  theme_minimal() +
  labs(title = "Observed vs. Predicted Wild Fish Caught (million tons) 1950-2016",
       x = "Year", 
       y = "Wild Caught Fish (million tons)") +
  scale_colour_manual(name="",
                    values = c(Predicted="slateblue4", Observed="gray62")) +
  theme(legend.position = c(0.28, 0.82),
        legend.text = element_text(size = 12)) 

```


