---
title: 'Agglomerative hierarchical clustering'
author: "Vanessa Rathbone"
date: "3/2/2021"
output:
    html_document:
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(palmerpenguins)

# Packages for cluster analysis: 
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```

Use hierarchical clustering by complete linkage to create a dendrogram showing multivariate clustering for water chemistry by site.

**Data Source:** Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. [Environmental Data Initiative.](https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174)

## Read in the data and prepare for the dendrogram
```{r, include = FALSE}
# To perform hierarchical clustering by site, you’ll want to make a data frame that has a single summary row per site (e.g. based on means from all observations at that site), then calculate the euclidean distance before performing complete linkage agglomerative hierarchical clustering. 
# Data drive: https://drive.google.com/file/d/16rYLBi-CgvjcWhOsX1SLdD9HHUMP9m2l/view?usp=sharing

# Read in the data and wrangle it, one site per column, remove -999
stream_df <- read_csv("sbc_lter_registered_stream_chemistry.csv", na = c("-999", "NaN")) %>% # change -999 to NA
  clean_names() %>% 
  group_by(site_code) %>% 
  summarize_if(is.numeric, list(mean), na.rm = TRUE) %>% # Summarize for dendrogram
  drop_na(nh4_u_m:spec_cond_u_spercm) %>% # Drop NAs
  mutate(site_code = case_when(site_code == "AB00" ~ "Arroyo Burro",
                               site_code == "AT07" ~ "Atascadero Creek",
                               site_code == "BC02" ~ "Bell Canyon Creek",
                               site_code == "DV01" ~ "Devereaux Creek",
                               site_code == "GV01" ~ "Gaviota",
                               site_code == "HO00" ~ "Arroyo Hondo",
                               site_code == "MC00" ~ "Mission Creek - Montecito St",
                               site_code == "MC06" ~ "Mission Creek - Rocky Nook",
                               site_code == "ON02" ~ "San Onofre Creek",
                               site_code == "RG01" ~ "Refugio Creek",
                               site_code == "RS02" ~ "Rattlesnake Creek",
                               site_code == "SP02" ~ "San Pedro Creek",
                               site_code == "TO02" ~ "Tecolote Creek"))

# To remove NAs or -999's you can also do mutate(across(where(is.numeric)), na_if(., -999.0))

# Scale the data
stream_scaled <- stream_df %>% 
  select(2:11) %>% 
  scale()

# Update to add rownames (stream name) as the rownames in the stream_scaled df
rownames(stream_scaled) <- stream_df$site_code

# Compute dissimilarity values (Euclidean distances): aka find the euclidean distance to find the dissimilarity values
euc_distance <- dist(stream_scaled, method = "euclidean")

```

## Make a dendrogram with heirarchical clustering using complete linkage
```{r, include = FALSE, echo = FALSE}

# Perform hierarchical clustering by complete linkage with `stats::hclust()`
hc_complete <- hclust(euc_distance, method = "complete")

# Plot it (base plot):
plot(hc_complete, cex = 0.6, hang = -1)

```

```{r, echo = FALSE}
# Make a dendrogram
ggdendrogram(hc_complete, 
             rotate = TRUE) +
  theme_minimal() +
  labs(title = "Heirarchical clustering using complete linkage", 
       subtitle = "Water chemistry by site", 
       x = "Site location", y = "") +
  theme(axis.ticks = element_blank()) +
  scale_y_continuous(breaks = NULL)

```

