---
title: 'Task 1: Agglomerative hierarchical clustering'
author: "Vanessa Rathbone"
date: "3/2/2021"
output: html_document
---

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(here)
library(janitor)
library(palmerpenguins)

# Packages for cluster analysis: 
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```

For Task 1, use hierarchical clustering by complete linkage to create a dendrogram showing multivariate clustering for water chemistry by site, saved as lastname_firstname_a4_task1.html. 

To perform hierarchical clustering by site, you’ll want to make a data frame that has a single summary row per site (e.g. based on means from all observations at that site), then calculate the euclidean distance before performing complete linkage agglomerative hierarchical clustering. 

Link to the data: https://drive.google.com/file/d/16rYLBi-CgvjcWhOsX1SLdD9HHUMP9m2l/view?usp=sharing

DATA & METADATA SOURCE: 
SBC LTER: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000
Creators: Santa Barbara Coastal LTER, & Melack, John M
Citation: Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 

```{r}
# Read in the data and wrangle it

chem_main <- 


```

#### Exploratory visualization

First, do some exploratory data visualization, mapping species onto point color. Does it look like there is an opportunity to cluster by species? 

```{r}

# Bill length versus depth exploratory plot:
ggplot(penguins) +
  geom_point(aes(x = bill_length_mm, 
                 y = bill_depth_mm, 
                 color = species,
                 shape = sex),
             size = 3,
             alpha = 0.7) +
  scale_color_manual(values = c("orange","cyan4","darkmagenta"))

# Flipper length versus body mass exploratory plot: 
ggplot(penguins) +
  geom_point(aes(x = flipper_length_mm, 
                 y = body_mass_g, 
                 color = species,
                 shape = sex),
             size = 3,
             alpha = 0.7) +
  scale_color_manual(values = c("orange","cyan4","darkmagenta"))
```

